---
title: "Lecture 3"
author: "Eryn McFarlane"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


We're going to simulate some data that can be used in each of the following tests.

The response variable is normally distributed, and then we're going to back calculate some predictor variables.

QUESTION: Describe the trait that you're simulating. This can be a real trait that you are working on, or hope to work on, or it can be completely made up. 
Pay special attention to the N that you're expecting, the mean, and the sd around the mean. 

```{r simulate trait of interest}
rnorm(1000, 0, 10)->trait_of_interest ### change this to be a trait that you're actually interested in, with an appropriate distribution! 

plot(density(trait_of_interest))
min(trait_of_interest)
max(trait_of_interest)

```

QUESTION: Describe the predictor variable. What does this mean biologically to your trait of interest. How did you decide on the numbers in yes on line 33?

```{r simulate predictor variable for a t test}

yes<-sample(trait_of_interest, 500, replace=FALSE, prob = ifelse(trait_of_interest>0, 0.95, 0.15)) #### play with this line! Is the test statistically significant. When is it not?
predictor_t.test<-(trait_of_interest %in% yes)

cbind.data.frame(trait_of_interest, predictor_t.test)->data

mean(data[which(data$predictor_t.test==TRUE),1])
mean(data[which(data$predictor_t.test==FALSE),1])
t.test(trait_of_interest~predictor_t.test, data=data) ### this does a two sample t-test. What would a one sample t test be testing? How would you do that?

### plots our two samples for distribution
plot(density(data[which(data$predictor_t.test==FALSE),1]), col="red", main="Two sample t test")
lines(density(data[which(data$predictor_t.test==TRUE),1]), ylim=c(0, 0.1), xlim=c(-20,20), main="Two Sample T test")

###plot one sample distribution
plot(density(data$trait_of_interest), col="red", main="One sample t test")
t.test(data$trait_of_interest) ### what is this test doing?


```


Next we're going to move to Anova. So, the first thing we'll do is break our response variable (same one!) into 5 different categories, just as we did for the t-tests. 

QUESTION: Describe the predictor variable. What does this mean biologically to your trait of interest. How did you decide on the numbers in lines 60, 61, 62, 63?

```{r Anova}

test1<-sample(trait_of_interest, 200, replace=FALSE, prob = ifelse(trait_of_interest>7, 0.95, 0.15))
test2<-sample(trait_of_interest[which(trait_of_interest %in% test1 == FALSE)], 200, replace=FALSE, prob = ifelse(trait_of_interest[which(trait_of_interest %in% test1 == FALSE)] >4, 0.95, 0.15))
test3<-sample(trait_of_interest[which(trait_of_interest %in% test2 == FALSE | trait_of_interest %in% test1 ==FALSE)], 200, replace=FALSE, prob = ifelse(trait_of_interest[which(trait_of_interest %in% test2 == FALSE | trait_of_interest %in% test1 == FALSE)] >0, 0.95, 0.15))
test4<-sample(trait_of_interest[which(trait_of_interest %in% test2 == FALSE | trait_of_interest %in% test1 ==FALSE | trait_of_interest %in% test3 == FALSE)], 200, replace=FALSE, prob = ifelse(trait_of_interest[which(trait_of_interest %in% test2 == FALSE | trait_of_interest %in% test1 ==FALSE | trait_of_interest %in% test3 == FALSE)] >-4, 0.95, 0.15))
test5<-trait_of_interest[which(trait_of_interest %in% test1 == FALSE| trait_of_interest %in% test2 == FALSE | trait_of_interest %in% test3 == FALSE | trait_of_interest %in% test4 == FALSE)]


plot(density(test1), ylim=c(0, 0.1), main="Anovas")
lines(density(test2), col="red")
lines(density(test3), col="blue")
lines(density(test4), col="purple")
lines(density(test5), col="yellow")


anova_predictor<-data.frame(ifelse(trait_of_interest %in% test1 == TRUE, "group1", ifelse(trait_of_interest %in% test2 == TRUE, "group2", ifelse(trait_of_interest %in% test3 == TRUE, "group3", ifelse(trait_of_interest %in% test4 == TRUE, "group4", "group5")))))

data2<-cbind.data.frame(data, anova_predictor)
names(data2)<-c("trait_of_interest", "predictor_t.test", "anova_predictor")


anova(aov(trait_of_interest~anova_predictor, data=data2)) ### what does this do?

anova(lm(trait_of_interest~anova_predictor, data=data2)) ### what does this do? 
summary(aov(trait_of_interest~anova_predictor, data=data2)) ### what does this do? What do you notice about the last three tests?

### what information is missing here that you wished you had to understand your study better?
analysis_of_variance<-aov(trait_of_interest~anova_predictor, data=data2) ##name the model to keep it for downstream
TukeyHSD(analysis_of_variance, conf.level = 0.95) ### what does this do, and where are the differences?

```


Moving on to linear regression - we've already talked about this, so I'm going to go quickly. 

Again, our simulations aren't to be actually causal, I'm simulating predictor variables to fit a response I've already made. Normally, we would simulate the response variables from more thoughtful predictor variables.

QUESTION:what is the difference between the assumed distributions for the prior predictor variables, and this one?
QUESTION: Describe the predictor variable. What does this mean biologically to your trait of interest. How did you decide on the numbers in line 101?

```{r Linear Regression}

linear_regression_predictor<-0.5*trait_of_interest+rnorm(1000, 0, 4) ### change these numbers!! Remember that this is backwards from how we did this on day 1, so the slope should go the other way!
data3<-cbind(data2, linear_regression_predictor)

lm<-lm(trait_of_interest~linear_regression_predictor, data=data3)
summary(lm)### what is the output here? What are we interested in understanding from this model? How do we get m? How do we get the intercept?

eq = paste0("y = ", round(lm$coefficients[2],1), "*x", "+",round(lm$coefficients[1],1), ", R^2=", round(summary(lm)$adj.r.squared, digits=3))
plot(data3$linear_regression_predictor, data3$trait_of_interest, col="red", main=eq)
abline(lm, col="black")


```