---
title: "Lecture 3"
author: "Eryn McFarlane"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

We're going to simulate some data that can be used in each of the following tests.

The response variable is normally distributed, and then we're going to back calculate some predictor variables.



```{r simulate trait of interest}
rnorm(1000, 0, 10)->trait_of_interest ### change this to be a trait that you're actually interested in, with an appropriate distribution! 

plot(density(trait_of_interest))
min(trait_of_interest)
max(trait_of_interest)

```

Describe the trait that you're simulating. This can be a real trait that you are working on, or hope to work on, or it can be completely made up. 
Pay special attention to the N that you're expecting, the mean, and the sd around the mean. 

```{r simulate predictor variable for a t test}

yes<-sample(trait_of_interest, 500, replace=FALSE, prob = ifelse(trait_of_interest>0, 0.95, 0.15)) #### play with this line! Is the test statistically significant. When is it not?
predictor_t.test<-(trait_of_interest %in% yes)

cbind.data.frame(trait_of_interest, predictor_t.test)->data

mean(data[which(data$predictor_t.test==TRUE),1])
mean(data[which(data$predictor_t.test==FALSE),1])
t.test(trait_of_interest~predictor_t.test, data=data)

```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
