---
title: "Lecture 5 Logistic, Poisson and Negative Binomial Regressions"
author: "Eryn McFarlane"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

We're going to simulate some predictor variables for each logistic regression, poisson regression and negative bionomial regression, and then analyze each. 

## Logistic Regression

We're going to have one response variable and two predictor variables (one which is yes or no, the other which is continuous).

### QUESTION: What are you imagining your response variable to be? What are your two predictor variables?

```{r logistic regression}
N=1000
predictor_cont<-rnorm(N,0,1) 
predictor_binary<-rbinom(N,1,0.5)

y<-1.5*predictor_cont+2.5*predictor_binary+rnorm(N, 0, 0.01) ### Remember that to make this do-able in a linear framework, like this, we're assuming this is on a log scale!

### We're now going to draw a probability distribution from y, assuming that y is on a logit scale

y_prob<-plogis(y)

###Question, tell me about the distribution of y_prob?!

### we can now draw from the probability distribution (so that all of our responses are 0,1)

y_dummy = rbinom(n = N, size = 1, prob = y_prob)


data<-cbind.data.frame(1:N, y_prob, y, predictor_cont, predictor_binary)

log_reg_model<-glm(y_dummy~predictor_binary+predictor_cont, data=data, family='binomial')

### if you have loaded and installed lmerTest, it will give you p-values in your summary, if not, it won't
summary(log_reg_model)

### From here, you need to back transform to get the estimates on the y scale, instead of the log(y) scale

exp(coef(log_reg_model)) ### these are on the 'regular' scale, not the log scale
exp(confint(log_reg_model))

```

### Question: What did you expect the effect sizes of the predictors to be, based on what you simulated (hint, you will need to transform from line 25)?

Let's simulate some Poisson data!

## Poisson Regression

This is mostly borrowed from [Simulate Simulate Simulate 3!](https://aosmith.rbind.io/2018/07/18/simulate-poisson-edition/)

### Question: What is your imagined response variable here? What is your imagined predictor?
### Question: What happens if you add more than one predictor (i.e. x2)?

```{r Poisson Regression}

x1 = rnorm(1000,1,0.1) ### here there is one predictors, this a good place to imagine some biology
summary(x1)

### Question, 

lambda = exp(0.5 + 2.5*x1) ### this is where I change the effect sizes for each predictor

###remember that this means that log(lambda) is a linear model!!
### here I'm simulating three betas against my three predictor variables to make my lambda
### The step above simulates the mean of each value of the response variable. lambda values are continuous, not discrete counts.
## λ is the unobserved true mean of the Poisson distribution for the t-th observation

head(lambda)

y = rpois(1000, lambda = lambda) ### error is added in this step, remembering that the mean and the variance are the same!
### remember that y is count data!
head(y)

plot(x1, log(y)) ### note the residuals - are they normally distributed? Is that okay?
plot(x1,y)

data_poisson<-cbind.data.frame(y, x1)

poisson_model<-glm(y ~ x1, family="poisson", data=data_poisson)
summary(poisson_model)

###back transform the coefficients 

exp(coef(log_reg_model)) ### these are on the 'regular' scale, not the log scale
exp(confint(log_reg_model))
## The coefficient for x1 is XX (FILL ME IN!) This means that the expected log count for a one-unit increase in x1 is XX. This means that the count difference (biological scale) for a one unit increase in x1 is XX. 
```




