---
title: "Lecture 6 - Mixed effects models"
author: "Eryn McFarlane"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(lme4)
library(lmerTest)
#install.packages("MuMIn")
library(MuMIn)
```

## Simulate some predictor variables

We're going to simulate three predictor variables and one random effect (nesting variable).
The three predictor variables are all going to affect our response variable. The random effect is just going to mess with this a bit

This nesting variable is going to 1) change the intercept for individuals within that group, and 2) change the error depending on who's in that group

```{r simulation}
a<-rnorm(1000, 10, 2) ### change these numbers, change these terms to something meaningful!
b<-rnorm(1000, 0, 1)
c<-rnorm(1000, 5, 1)

g<-sample(1:3, 1000, replace=TRUE) 
intercept<-ifelse(g==1, 2, ifelse(g==2, -2, 0))
error<-ifelse(g==1, 0.1, ifelse(g==2, 0.5, 1.5))
g<-as.factor(g)
response<-intercept+0.5*a+0.2*b+0.1*c+rnorm(1000, mean=0, sd=error)
data<-cbind.data.frame(response, a, b, c, g)
```

Let's look at the plots of the data, and then the data against (at least one) of the predictor variables. Feel free to plot separately against each predictor variable, or, plot partial residuals (remember Lecture 4!
)
```{r plots}
hist(response) ### lovely and normally distributed

ggplot(data, aes(y=response, x=a))+geom_point()+geom_smooth(method=lm)

ggplot(data, aes(y=response, x=a, colour=g))+geom_point()+geom_smooth(method=lm)

```

QUESTION: Describe the difference between the two scatter plots. What does adding the different colour lines show you?

QUESTION: Describe the residuals in both plots.

Let's do some analyses on this to take a look at the differences. 
```{r analysis}


model1<-lm(response~a+b+c, data=data)
summary(model1)
anova(model1)

model2<-lmer(response~a+b+c+(1|g), data=data)
summary(model2)
anova(model2)

```

```{r interpretation}
## R2 for each model
summary(model1)$adj.r.squared

r.squaredGLMM(model2) ### this gives two R2, what are they both?

##intercepts for each model
 summary(model1)$coef[1,1]
 ranef(model2)$g

 ### coefficents for each model
model1$coefficients
 
summary(model2)$coefficients[, "Estimate"]
 
```

QUESTION: Compare and contrast the R2, intercepts and estimates from each model. Which is closer to what you simulated? Why is this?
